---
title: "Bioconductor Proteomics Agent -- Analysis Report"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S %Z')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    self_contained: true
params:
  objective: "Not specified"
  file_meta: NULL
  qc_metrics: NULL
  msquality_metrics: NULL
  processing_comparison: NULL
  identification_summary: NULL
  quantification_summary: NULL
  msstats_summary: NULL
  plan_text: NULL
  trace_df: NULL
  warnings: NULL
  session_provenance: NULL
  ai_narrative: NULL
  llm_mode: "mock"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
`%||%` <- function(a, b) if (is.null(a) || length(a) == 0) b else a
```

> **Research demonstration only. Not intended for clinical diagnosis or clinical decision-making.**

## Analysis Objective

`r htmltools::htmlEscape(params$objective %||% "Not specified")`

## Source File Metadata

```{r file-meta}
if (!is.null(params$file_meta)) {
  knitr::kable(data.frame(
    field = names(params$file_meta),
    value = vapply(params$file_meta, function(x) paste(x, collapse = ", "), character(1))
  ))
} else {
  cat("No source file metadata was recorded for this run.")
}
```

## Experiment Overview & QC Metrics

```{r qc-table}
if (!is.null(params$qc_metrics)) {
  qc <- params$qc_metrics
  flat <- data.frame(
    metric = c("Total spectra", "MS1 spectra", "MS2 spectra", "MS2/MS1 ratio",
               "RT range (s)", "Missing precursor (MS2)", "Empty spectra"),
    value = c(qc$n_spectra, qc$n_ms1, qc$n_ms2,
              sprintf("%.3f", qc$ms2_ms1_ratio %||% NA),
              paste(round(qc$rt_range_s %||% c(NA, NA), 1), collapse = " - "),
              qc$n_missing_precursor, qc$n_empty_spectra)
  )
  knitr::kable(flat)
} else {
  cat("QC metrics were not computed for this run.")
}
```

### QC Warnings

```{r qc-warnings, results='asis'}
if (!is.null(params$qc_metrics) && length(params$qc_metrics$warnings) > 0) {
  cat(paste0("- ", htmltools::htmlEscape(params$qc_metrics$warnings), collapse = "\n"))
} else {
  cat("No QC warnings were flagged.")
}
```

### Standardized QC Metrics (MsQuality)

```{r msquality-table}
if (!is.null(params$msquality_metrics) && length(params$msquality_metrics) > 0) {
  m <- params$msquality_metrics
  knitr::kable(data.frame(
    metric = names(m),
    value = vapply(m, function(x) paste(format(x, digits = 4), collapse = ", "), character(1))
  ))
} else {
  cat("Standardized (MsQuality) QC metrics were not computed for this run.")
}
```

## Processing: Before / After Comparison

```{r processing-comparison}
if (!is.null(params$processing_comparison)) {
  cmp <- params$processing_comparison
  flat <- data.frame(
    metric = names(unlist(cmp)),
    value = unname(unlist(cmp))
  )
  knitr::kable(flat)
} else {
  cat("No spectrum processing stage was recorded for this run.")
}
```

## Identification Summary

```{r ident-summary}
if (!is.null(params$identification_summary)) {
  knitr::kable(data.frame(
    metric = names(params$identification_summary),
    value = vapply(params$identification_summary, function(x) paste(x, collapse = ", "), character(1))
  ))
} else {
  cat("No PSM identification table was analyzed in this run.")
}
```

## Quantification Summary

```{r quant-summary}
if (!is.null(params$quantification_summary)) {
  knitr::kable(data.frame(
    metric = names(params$quantification_summary),
    value = vapply(params$quantification_summary, function(x) paste(x, collapse = ", "), character(1))
  ))
} else {
  cat("No quantitative abundance table was analyzed in this run.")
}
```

### Statistical Comparison (MSstats)

```{r msstats-summary}
if (!is.null(params$msstats_summary)) {
  knitr::kable(data.frame(
    metric = names(params$msstats_summary),
    value = vapply(params$msstats_summary, function(x) paste(x, collapse = ", "), character(1))
  ))
} else {
  cat("No MSstats comparison was run for this session.")
}
```

## Agent Plan

```{r plan-text, results='asis'}
if (!is.null(params$plan_text)) {
  cat(paste0("<pre>", htmltools::htmlEscape(params$plan_text), "</pre>"))
} else {
  cat("No plan was recorded for this run.")
}
```

## Tool Calls & Agent Trace

```{r trace-table}
if (!is.null(params$trace_df) && nrow(params$trace_df) > 0) {
  knitr::kable(params$trace_df[, c("timestamp", "agent", "tool", "status", "duration_s", "output_id")])
} else {
  cat("No agent trace was recorded for this run.")
}
```

## Warnings

```{r warnings-list, results='asis'}
if (!is.null(params$warnings) && length(params$warnings) > 0) {
  cat(paste0("- ", htmltools::htmlEscape(params$warnings), collapse = "\n"))
} else {
  cat("No warnings were raised during this run.")
}
```

## AI-Assisted Interpretation

**AI-assisted interpretation generated using Claude**

```{r ai-narrative, results='asis'}
if (identical(params$llm_mode, "mock")) {
  cat("_Mock mode was active for this run; the text below was produced by the deterministic mock backend, not by Claude._\n\n")
}
if (!is.null(params$ai_narrative) && nzchar(params$ai_narrative)) {
  cat(htmltools::htmlEscape(params$ai_narrative))
} else {
  cat("No AI interpretation was generated for this run.")
}
```

*The section above is AI-generated narrative interpretation. All figures, tables, and metrics elsewhere in this report were produced by deterministic R/Bioconductor functions, not by the language model.*

## Limitations

- This tool is a research and workshop demonstration. It is not validated for clinical or diagnostic use.
- Proprietary vendor formats (Thermo RAW, Bruker, SCIEX, Waters) are not read directly; convert to mzML first (e.g. with ProteoWizard `msconvert`).
- Statistical comparisons are only performed when minimum replicate requirements are met; results with small sample sizes should be interpreted cautiously.
- AI-assisted interpretation may be incomplete or imprecise; it is grounded in, but not a substitute for, the computed metrics above.

## Package & Session Provenance

```{r provenance}
sp <- params$session_provenance
if (!is.null(sp)) {
  knitr::kable(data.frame(
    field = c("Generated at", "Git commit SHA", "Bioconductor version", "R version"),
    value = c(sp$generated_at, sp$git_commit_sha, sp$bioc_version %||% "unknown", sp$r_version)
  ))
  cat("\n\n### Key package versions\n\n")
  knitr::kable(data.frame(package = names(sp$package_versions), version = unlist(sp$package_versions)))
} else {
  cat("Session provenance was not recorded.")
}
```

```{r session-info}
if (!is.null(sp$session_info)) {
  cat(paste0("<details><summary>Full R session info</summary><pre>",
             htmltools::htmlEscape(sp$session_info), "</pre></details>"))
}
```

